---
title: "Visualization of EEG Signals in MEDA"
author: "Ronak Mehta and Yujia Liu"
date: "September 25, 2017"
output: pdf_document
---

Below is a set of visualizations using the MEDA library from NeuroData. Our data is a set of various
observations (about 200,000) from a 128-channel EEG test.

```{r}
subject_file <- "C:/Users/Ronak Mehta/Desktop/NeuroData/NDARAA075AMK_RestingState_data.csv"
first_obs_no <- 1
last_obs_no <- 100001
sep <- 50
```

## Setting Variables and Ingesting Data

The following selects the columns to take from the ingested CSV. In this case, we take the first 100000
samples, and use every 50th data point to make these visualizations. 

```{r}
library(data.table, quietly = TRUE)
library(meda, quietly = TRUE)

# Set the columns (observations) we want to take.
columns <- seq(first_obs_no, last_obs_no, sep)

# Set column names to the electrodes.
col_names <- rep("E", 128)
for (i in 1:length(col_names)) { col_names[i] <- paste(col_names[i], as.character(i), sep = "") }

# Injest CSV of subject's data.
dat <- transpose(fread(input = subject_file, 
                                sep = ",", 
                                nrows = 128,
                                header = FALSE,
                                showProgress = TRUE,
                                select = columns,
                                data.table = FALSE))
colnames(dat) <- col_names
```

## Plot Exploratory Analyses

Each plot is shown below.

### Location Plots

```{r}
plot(mlocation(dat))
```

### Heat Map
```{r}
print(plot(d1heat(dat)))
```

### Outlier Plot
```{r}
print(plot(outliers(dat)))
```

### Correlation Matrix
```{r}
plot(medacor(dat))
```

### Cumulative Variance Plot
```{r}
print(plot(cumvar(dat)))
```

### Fit Clusters with GMM
```{r}
h <- hmc(dat)
```

### Dendogram
```{r}
plotDend(h)
```

### Stacked Cluster Means Plot
```{r}
print(stackM(h, centered = TRUE))
```

### Cluster Means
```{r}
clusterMeans(h)
```